---
title: "daily_longfin_salvage"
format: html
editor: visual
---

```{r setup, include=FALSE}
library(here)
library(dplyr)
library(readr)
library(ggplot2)
library(janitor)
library(lubridate)
library(viridis)
library(readxl)
```

## Notes on Processed Data

-   Removed any row without a count, length, sample fraction
-   Filtered to 2014-2025
-   Assigned Life Stage (Adult \> 60mm; Juvenile \<= 60 mm)
-   Cumulative Salvage represents fish count/sample fraction

```{r}
lfs_data <- read_excel("WY2025_smelt_catch.xlsx", sheet = 2)  %>%
  clean_names()

df_2025 <- lfs_data %>%
  filter(salvage_sum>0)%>%
  rename(expanded = exp_salv,
         length = fork_length,
         lifestage = life_stage) %>%
  mutate(lifestage = if_else(length>60, "Adult", if_else(length>=19, "Juvenile", "Larva"))) %>%
  mutate(wy = "2025",
         date = ymd(date),
         month = month(date),
         date2 = case_when(month>=10 ~ ymd(paste0(1980, "-", month, "-", day(date))),
                           month<10 ~ ymd(paste0(1981,"-", month, "-", day(date) )))) %>%
  mutate(facility = if_else(station_code == "SFF", "SWP", "CVP")) %>%
  group_by(lifestage) %>%
  mutate(cumulative_salvage = cumsum(expanded)) %>%
  ungroup() %>%
  select(facility, date, length, wy, month, date2, lifestage, expanded, cumulative_salvage)
```

```{r, include = FALSE}
lfs_salvage0 <- read_csv(here("data_raw/lfs_loss.csv")) 
lfs_salvage <- lfs_salvage0 %>%
  clean_names() %>%
  filter(!is.na(facility),
         !is.na(nfish),
         !is.na(length)) %>%
  mutate(sample_time = mdy_hm(sample_time),
         year = year(sample_time),
         month = month(sample_time),
         date = date(sample_time),
         wy = if_else(month>=10, year+1, year)) %>%
  # filter(wy > 2013) %>%
  mutate(wy = as.factor(wy)) %>%
  mutate(date2 = case_when(month>=10 ~ ymd(paste0(1980, "-", month, "-", day(date))),
                           month<10 ~ ymd(paste0(1981,"-", month, "-", day(date) )))) %>%
  mutate(lifestage = if_else(length>60, "Adult", if_else(length>=19, "Juvenile", "Larva"))) %>%
  mutate(expanded = nfish/sample_fraction)%>%
  filter(!is.na(expanded)) %>%
  group_by(wy, lifestage) %>%
  arrange(date) %>%
  mutate(cumulative_salvage = cumsum(expanded)) %>%
  ungroup() %>%
  select(facility, date, length, wy, month, date2, lifestage, expanded, cumulative_salvage)
```

```{r, include = FALSE}
## Combine data frames
lfs_salvage_all <- bind_rows(lfs_salvage, df_2025) %>%
  mutate(current_year = if_else(wy == 2025, "Y", "N"))

lfs_salvage_juv <- lfs_salvage_all %>% 
  filter(lifestage == "Juvenile") 

lfs_salvage_adult <- lfs_salvage_all %>% 
  filter(lifestage == "Adult") 


```


## Juvenile Plot
```{r longfin-plot, echo = FALSE, fig.cap = "Longfin Smelt Salvage, WY 2014-2025. Data from SacPAS/CDFW.", fig.width = 9, fig.height = 6}
## Plot results
scaledf <- lfs_salvage_all[which(!duplicated(lfs_salvage_all$lifestage)),]

currentj = max(df_2025%>% 
                 filter(lifestage == "Juvenile") %>% 
                 select(cumulative_salvage))

(plot_juv_lfs_salvage <- ggplot() +
  geom_point(data = scaledf, alpha = 0, aes(x = date2, y = cumulative_salvage)) + 
  geom_point(data=lfs_salvage_juv, aes(x = date2, y = cumulative_salvage, color = wy, shape = wy, group = wy, size = current_year, alpha = current_year)) +
  geom_line(data=lfs_salvage_juv, aes(x = date2, y = cumulative_salvage, color = wy, group = wy, alpha = current_year)) + 
    geom_hline(yintercept = 618, linetype = "dashed")+
    geom_hline(yintercept = 927, linetype = "dashed")+
    annotate(geom = "text", x = ymd("1981-03-10"), y = 450, label = "50% Threshold: 620")+ 
    annotate(geom = "text", x = ymd("1981-03-10"), y = 1100, label = "75% Threshold: 927")+ 
  annotate(geom = "text", x = ymd("1981-03-20"), y = 6500,size = 4, label = paste0("Current 2025 \nCumulative Salvage:", currentj,"\n(as of ", format(today(), "%m/%d"),")"))+ 
    annotate("rect", xmin = ymd("1981-03-04"), xmax = ymd("1981-04-05"), ymin = 5850, ymax = 7150, fill = NA, color = "black")+
  labs(x = "Date", color = "Water Year", shape = "Water Year", y = "Cumulative LFS Salvage\n(expanded)", title = "Juvenile Longfin Smelt Salvage\nWY 2014-2025") +
  scale_color_viridis(discrete = TRUE, option = "turbo")+
  scale_alpha_manual(values = c(0.3, 1)) + 
  scale_size_manual(values = c(1.6, 3)) +
  scale_shape_manual(values = c(25, 4, 1, 23, 15, 7, 23, 3, 19, 0, 8))+
  scale_x_date(date_labels = "%b") +
  theme_bw() +
  guides(size = "none", alpha = "none") +
  theme(strip.text = element_text(size = 12),
        axis.text = element_text(size = 12),
        legend.text = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.title = element_text(size = 13),
        legend.title = element_text(size = 13),
        plot.title = element_text(hjust = 0.5)))
```

```{r, include = FALSE}
png(paste0("lfs_juv_daily_salvage_plot_", today(), ".png"), res = 300, width = 8, height = 5, units = "in")
plot_juv_lfs_salvage
dev.off()
```

## Adult Plot
```{r}
currenta = max(df_2025%>% 
                 filter(lifestage == "Adult") %>% 
                 select(cumulative_salvage))

lfs_salvage_adult <- lfs_salvage_adult %>% 
filter(date2<ymd("1981-07-01"))

(plot_adult_lfs_salvage <- ggplot() +
  geom_point(data = scaledf, alpha = 0, aes(x = date2, y = cumulative_salvage)) + 
  geom_point(data=lfs_salvage_adult, aes(x = date2, y = cumulative_salvage, color = wy, shape = wy, group = wy, size = current_year)) +
  geom_line(data=lfs_salvage_adult, aes(x = date2, y = cumulative_salvage, color = wy, group = wy)) + 
    # geom_hline(yintercept = 181.2, linetype = "dashed")+
    # annotate(geom = "text", x = ymd("1980-12-25"), y = 175, label = "Threshold: 181.2")+ 
  # annotate(geom = "text", x = ymd("1981-02-08"), y = 150,size = 4, label = paste0("2025 Cumulative Salvage: ", currenta,"\n(as of ", format(today(), "%m/%d"),")"))+ 
  # annotate("rect", xmin = ymd("1981-01-21"), xmax = ymd("1981-02-26"), ymin = 135, ymax = 165, fill = NA, color = "black")+
  labs(x = "Date", color = "Water Year", shape = "Water Year", y = "Cumulative LFS Salvage\n(expanded)", title = "Adult Longfin Smelt Salvage\nWY 1995-2025") +
  scale_color_viridis(discrete = TRUE, option = "turbo")+
  scale_alpha_manual(values = c(0.7, 1)) + 
  scale_size_manual(values = c(1.6, 3)) +
  scale_shape_manual(values = c(15, 8, 23, 16, 7, 23, 3, 0,
                                12, 17, 9, 13, 4, 6, 19, 18))+
  scale_x_date(date_labels = "%b") +
  theme_bw() +
  guides(size = "none", alpha = "none") +
  theme(strip.text = element_text(size = 12),
        axis.text = element_text(size = 12),
        legend.text = element_text(size = 12),
        axis.title.x = element_blank(),
        axis.title = element_text(size = 13),
        legend.title = element_text(size = 13),
        plot.title = element_text(hjust = 0.5)))
```

```{r, include = FALSE}
png(paste0("lfs_adult_daily_salvage_plot_", today(), ".png"), res = 300, width = 8, height = 5, units = "in")
plot_adult_lfs_salvage
dev.off()
```
